---
title: "Data Science - Avocado Prices"
author: "Jens Mattsson"
date: "2020 M06 30"
output:
  pdf_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###R tips and tricks

Cursor inside a command and press f1 will open the help window.
cursor inside a chunk: Run current line: ctrl+enter. Run whole chunk: ctrl+shit+enter.
When saving code and uploading to github: Git, Diff, click checkbox, write some commit command, commit, then pull, then push.
remove()


```{r Load packages and data set, echo=FALSE, include=FALSE}
packages = c("stargazer", "ggplot2", "lubridate", "dplyr")
lapply(packages, library, character.only = TRUE)

avocado = read.csv("avocado.csv", header = T, sep = ",")
```



##Inspect, plot, and check descriptive statistics

```{r fix the Date variable, echo=FALSE, include=FALSE}
is.Date(avocado$Date)
avocado$Date = as.Date(avocado$Date)
is.Date(avocado$Date)
str(avocado)
avocado$AveragePrice = as.numeric(avocado$AveragePrice)

```

```{r rename variables, echo=FALSE, include=FALSE}
names(avocado)[5] = "Small"
names(avocado)[6] = "Large"
names(avocado)[7] = "XL"
```


```{r echo=FALSE}
names(avocado)
ggplot(avocado) + geom_point(aes(y = AveragePrice, x = Date), size = 0.2) + scale_x_date(date_breaks = "1 year") + geom_smooth(aes(y = AveragePrice, x = Date))



'col = colnames(avocado)
for (i in col) {
  ggplot(avocado) + geom_point(aes(y = AveragePrice, x = i), size = 0.2)
}
for (i in colnames(avocado)[2][4:14]){
ggplot(avocado) + geom_point(aes(y = AveragePrice, x = i), size = 0.2)
}
for (i in col){
plot(avocado$AveragePrice,avocado$i)
  }'
```


```{r}
length(unique(avocado$region))
```

```{r echo=F, }
ggplot(avocado,  aes(y = AveragePrice, x = Date)) +  geom_point(color="blue", size = 0.2)

```


%%%%%%%%From Testing.Rmd


#Let´s plot the average price as for each period

#The price looks pretty messy right? So let´s use Pontus coola confidence interval grej

```{r, echo=F}
price_plot = ggplot(avocado) +
  geom_point(aes(y = AveragePrice, x = Date), size = 0.1) + 
  scale_x_date(date_breaks = "6 months")
price_plot
price_plot +  geom_smooth(aes(y = AveragePrice, x = Date))
```
As we can see, there appears to be some cyclical pattern that is further enhanced by time. Could data disaggregate  based on months be reasonable?
Varaiance might increase aswell, may use a GARCH similar model

```{r cool histogram, echo=F}
#avocado_ps = pdata.frame(avocado, index = c("region", "Date"))
#model = plm(formula = AveragePrice ~ lag(AveragePrice, 1:5), subset(avocado_ps, type == "conventional"))
#summary(model)
ggplot(avocado, aes(x = AveragePrice, fill = type)) +
  geom_histogram()

#The following code does not produce what we want.
kolumner = colnames(avocado)
for(j in kolumner){
  print(ggplot(avocado, aes(y = AveragePrice, x = j)) + geom_point())
}
```



```{r geom_dotplot price month and geom_tile, echo=F}
#Making new variables. Sebbe did not notice that we already had a year variable :)
avocado$Year = format(as.Date(avocado$Date), "%Y")
avocado$Month = format(as.Date(avocado$Date), "%m")
#Gets the average and variance of each month for all years and states
avocado = avocado %>% 
  group_by(Month) %>%
  summarise(avg_M = mean(AveragePrice), var_M = var(AveragePrice)) %>% 
  merge(avocado, by = c("Month"))

#Gets the average and variance subsetted on each month and year. states are not differenteiated between
avocado = avocado %>% 
  group_by(Month, Year) %>%
  summarise(avg_M_Y = mean(AveragePrice), var_M_Y = var(AveragePrice)) %>% 
  merge(avocado, by = c("Month", "Year"))

#Average price per month
ggplot(data = avocado, aes(x = Month, y = avg_M)) +
  geom_point()

#Average price per month plus boxplot.
ggplot(data = avocado, aes(x = Month, y = AveragePrice)) +
  geom_boxplot() +
  geom_point(aes(y = avg_M), colour = "blue")

#Tile plot with Month, Year and Average price.
ggplot(avocado, aes(x = Month, y = Year, fill = avg_M_Y)) +
  geom_tile() +
  scale_fill_viridis_c()
```
      
